---
title: "Homework 2"
author: "Beibei Cao"
date: "2020-9-24"
output: github_document
---

This is my solution to Homework 2.

```{r setup, message = FALSE}
# load libraries that will be used in this project
library(tidyverse)
library(readxl)
```


## Problem 1

Read the Mr. Trashwheel dataset.

```{r}
trashwheel_df = 
  read_xlsx(
    "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", 
    sheet = "Mr. Trash Wheel",
  range = cell_cols("A:N")) %>% 
  janitor::clean_names() %>% 
  drop_na(dumpster) %>% 
  mutate(
    sports_balls = round(sports_balls),
    sports_balls = as.integer(sports_balls)
  )
```

Read precipitation data!

```{r}
precip_2018 = 
  read_excel(
    "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
    sheet = "2018 Precipitation",
    skip = 1
  ) %>% 
  janitor::clean_names() %>% 
  drop_na(month) %>% 
  mutate(year = 2018) %>% 
  relocate(year)

precip_2017 = 
  read_excel(
    "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
    sheet = "2017 Precipitation",
    skip = 1
  ) %>% 
  janitor::clean_names() %>% 
  drop_na(month) %>% 
  mutate(year = 2017) %>% 
  relocate(year)
```

Now combine the annual precipitation.

```{r}
month_df = 
  tibble(
    month = 1:12,
    month_name = month.name
  )

precip_df = 
  bind_rows(precip_2018, precip_2017)

left_join(precip_df, month_df, by = "month")
```

This dataset contains information from the Mr. Trashwheel trash collector in Baltimore, Maryland. As trash enters the inner harbor, the trashwheel collects that trash, and stores it in a dumpster. The dataset contains information on year, month, and trash collected, including some specific kinds of trash. There are a total of ``r nrow(trashwheel_df)`` rows in our final dataset. Additional date sheets include month precipitation data. 


## Problem 2
#### This problem focuses on NYC Transit data.

Read and clean the NYC Transit data.

```{r collapse = TRUE}
# read NYC csv
nyctransit_df = 
  read_csv(
    file = "./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>% 
  janitor::clean_names() 

# check columns  
colnames(nyctransit_df)

# select wanted columns
nyctransit_df = nyctransit_df[,c(2:18,20,23)]

# preview and check which column with character (YES vs NO)
str(nyctransit_df)
  
# convert entry variable (col 17-18, 'entry' and 'vending') from character (YES vs NO) to a logical variable
 for (cl in c(17, 18)){
  nyctransit_df[cl] = 
    ifelse(nyctransit_df[cl] == "YES", TRUE, FALSE)
}

# check result
str(nyctransit_df[17:18])
```

So far, we load the NYC Transit dataset, cleaned the names of columns, selected the desired columns and changed two column, `entry` and `vending`,  from `YES/NO` character variables to `TRUE/FALSE` logical variables. The dataset contained the following ``r ncol(nyctransit_df)`` variables: ``r names(nyctransit_df)``. There are now ``r nrow(nyctransit_df)`` rows in the dataset. 
